WORCS: A Workflow for Open Reproducible Code in Science

### 2020-10-31 paper/Manuscript_DS.Rmd
R3 C3: Git not useful

An integral part of WORCS is the *explicit* version control software Git
(www.git-scm.com). Git tracks
changes to files, and stores these changes when the user makes a "commit" (a 
snapshot of the version controlled files).
Git retains a complete log of all commits, and users can compare changes between
different commits, or go back to a previous version of the code (for example,
after making a mistake, or to replicate a previous version of the results).
A project version controlled with Git is called a "Repository", or Repo.
Git only version controls files explicitly committed by the user. Moreover, it
is possible to prevent files from being version controlled - which is useful for
privacy sensitive data.
A `.gitignore` file lists files that should not be version-controlled.
These files thus exist only on the user's private computer.

@@

A key consideration when developing a research project is what filetypes to use.
WORCS encourages the use of text-based files whenever possible,
instead of binary files.
Text-based files can be read by machines and humans alike.
Binary files, such as Word (`.docx`) or SPSS (`.sav`, `.spo`) files,
must be decoded first, often requiring commercial software.
Git is designed for use with text-based files, as it
logs changes to text-based files on a line-by-line basis.
Although binary files can be version controlled with Git, 
their change log is uninterpretable.
Binary files are also often 
larger than text-based files, which means they take up more space on cloud
hosting services. For these reasons, uploading large binary
files to remote repositories is frowned upon.

>>>>

An integral part of WORCS is the *explicit* version control software Git
(www.git-scm.com).
Git tracks changes to files on a line-by-line basis.
The user can sture these changes by making a "commit": a 
snapshot of the version controlled files.
Each "commit" can contain as many changes as desired;
for example, a whole new paragraph, or many small changes made to address a single reviewer comment.
Git retains a complete log of all commits, and users can compare changes between
different commits, or go back to a previous version of the code (for example,
after making a mistake, or to replicate a previous version of the results).
A project version controlled with Git is called a "Repository", or Repo.
Git only version controls files explicitly committed by the user. Moreover, it
is possible to prevent files from being version controlled - which is useful for
privacy sensitive data.
A `.gitignore` file lists files that should not be version-controlled.
These files thus exist only on the user's private computer.

@@

A key consideration when developing a research project is what filetypes to use.
WORCS encourages the use of text-based files whenever possible,
instead of binary files.
Text-based files can be read by machines and humans alike.
Binary files, such as Word (`.docx`) or SPSS (`.sav`, `.spo`) files,
must be decoded first, often requiring commercial software.
Although binary files can be version controlled with Git, 
their change log is uninterpretable.
Binary files are also often 
larger than text-based files, which means they take up more space on cloud
hosting services. For these reasons, uploading large binary
files to remote repositories is frowned upon.

### 2020-10-31 paper/reviews_ds.Rmd
R3 C3: Git not useful

\Assignment{ALL}
\WorkInProgress
\Hard

@@

Response: We agree with the Reviewer that the conceptual workflow as explained in the paper could be implemented in any software environment, including Jupyter. In fact, we acknowledge this in the paragraph on *Future developments*: *"although the workflow is currently implemented only in `R`, it is conceptually relevant for users of other statistical programming languages, most notably `Python` [@10.5555/1593511]. We welcome efforts to implement WORCS in other platforms."*.

The simple and straightforward response to the Reviewer's question is that we implemented this workflow in R because we are R developers. But in the paper, we provide some more detailed arguments to substantiate a preference for R over other environments:

@@

Response: Ideas?

>>>>

@@

Response: There are two points we wish to address here. First, we would like to clarify that this paper introduces WORCS, a conceptual workflow for open science research projects. When the Reviewer refers to "the tool" - we assume this to mean the `worcs` R-package, which offers convenience functions to help R-users follow the WORCS-workflow. This is clarified in the Abstract and in the paper, e.g.:

> Abstract: *This paper introduces the Workflow for Open Reproducible Code in Science (WORCS): A step-by-step procedure that researchers can follow to make a research project open and reproducible, accompanied by an open source software implementation that automates these steps.*

> Manuscript, paragraph titled **The R implementation of WORCS**: *WORCS is, first and foremost, a conceptual workflow that could be implemented in any software environment. As of this writing, the workflow has been implemented for `R` users in the package `worcs` [@worcspackage2020].*

Second, in response to the main comment, we agree with the Reviewer that the conceptual workflow as explained in the paper could be implemented in any software environment, including Jupyter. In fact, we acknowledge this in the paragraph on *Future developments*: *"although the workflow is currently implemented only in `R`, it is conceptually relevant for users of other statistical programming languages, most notably `Python` [@10.5555/1593511]. We welcome efforts to implement WORCS in other platforms."*.

To address the Reviewer's question directly: We implemented this workflow in R because we are R developers. But in the paper, we also provide some arguments that compel our preference for R and RStudio as a choice for statistical computing in open science:

@@

Response: Respectfully, we are not entirely sure what the criticism is here. First, there seems to be a misunderstanding about the scope of the paper, which is described in the Abstract: *"This paper introduces [a] step-by-step procedure that researchers can follow to make a research project open and reproducible, accompanied by an open source software implementation that automates these steps"*.

Our paper thus introduces a workflow for open science, based on objective criteria set out in the TOP-guidelines. Aside from the work in progress of our co-authors - which is focused on strict computational reproducibility and is compatible with the present workflow - this is one of the first efforts to translate the requirements of open science into a step-by-step workflow. When you search Google Scholar for "open science workflow", the first hit that actually describes a workflow is the present manuscript.

Second, the Reviewer claims that using e.g. Git does not add substantial value - but this claim is in disagreement with published literature that makes a strong case for the value of using Git for research [see two references cited in the paper, @ramGitCanFacilitate2013;  @blischakQuickIntroductionVersion2016].

Third, the Reviewer remarks that 
"git capture changes at the level of the line of code/text, whereas a researcher needs an abstraction that informs him/her on the coarse-grained elements [...] so as to have a clear idea of where s/he is within the process and help him/her make the decision of the next steps to undertake".
This comment suggests a fundamental misunderstanding of how Git works.
Although Git tracks file changes on a line-by-line basis,
any number of such changes can be committed together, thereby capturing the more "course-grained elements".
For example, in our response to these reviews, we have created one Git commit for each Reviewer comment, and these commits contain changes to the action letter, manuscript, and vignettes.

We do see the value in clarifying this point in the manuscript as well, and to address this comment, the manuscript now reads:

> Git tracks changes to files on a line-by-line basis. The user can sture these changes by making a "commit": a snapshot of the version controlled files. Each "commit" can contain as many changes as desired; for example, a whole new paragraph, or many small changes made to address a single reviewer comment.

\Assignment{CJ}
\Done
\Medium
